% SHARD-ID: AQM-02-LAB-LOG
% SHARD-TITLE: Lab Log
% SHARD-SUMMARY: Maintains a chronological record of infrastructure changes and sourced research steps.
% SHARD-SUMMARY: Later entries record sourced toric/CSS/Steane and arithmetic-field results.
% SHARD-KEYWORDS: lab log, session log, chronology, handoff

\section{Lab Log}

\subsection*{2026-05-24}

Initial lab-book scaffold created.

\begin{itemize}
  \item Established \texttt{report.tex} as the sharded LaTeX master.
  \item Added report source maps, convention and index files, source-manifest
  placeholders, run-bundle rules, and a tool runner.
  \item Recorded that no sources have been acquired and no scientific claims
  have been checked yet.
\end{itemize}

\subsection*{2026-05-24 Steane Clifford Ghost Gaussians}

The report now includes sourced toric/CSS/Steane stabilizer-supercharge
derivations.  The latest addition is
Section~\ref{sec:clifford-group-ghost-gaussian-theorem}: arbitrary Clifford
covariance is stated for signed Pauli stabilizer presentations, and
presentation changes are lifted by number-preserving fermionic Gaussian
operations on the ghosts.  The accompanying Steane run checks all standard
H/P/CNOT Clifford generators and the elementary GL6 ghost moves.

\subsection*{2026-05-24 Arithmetic Quantum Fields}

Section~\ref{sec:arithmetic-quantum-fields} adds the first explicit
arithmetic-quantum-field proposal.  A finite function space
\(E\leq\operatorname{Map}(X,V)\) receives a pointwise symplectic form; after
quotienting its radical, its Weyl-Heisenberg displacements define the field
observable algebra.  The run bundle checks finite-set, vector-space, and
finite-variety examples over \(F_3\).

\subsection*{2026-05-24 Projective Sheaf Fields}

Sections~\ref{sec:projective-sheaf-field-definitions} and
\ref{sec:projective-line-sheaf-field-calculation} add the first projective
version; Section~\ref{sec:projective-line-stalks} works out the rational-point
stalks; Section~\ref{sec:stalks-as-field-germs} records the working dictionary
``stalk equals local field germ, fiber equals sampled field value.''  Local
Stacks Project TeX sources now ground the definitions of
presheaf, sheaf, stalk, projective space, varieties, twists, and
\(H^0(\mathbf P^n,\mathcal O(d))\).  The run bundle checks
\(\mathbf P^1_{\mathbb F_3}\) with \(H^0(\mathcal O(d))\otimes\mathbb F_3^2\)
for \(d=0,\ldots,4\), including the \(d=1\) pairing radical, the \(d=4\)
evaluation kernel, the local germs whose residues give the evaluation rows,
and the full \(d=0,\ldots,4\) scalar residue table.

\subsection*{2026-05-25 Canonical Boson/Fermion Field Test}

Section~\ref{sec:canonical-boson-fermion-field-comparison} adds a sourced
finite-mode comparison with standard canonical fields.  The bosonic test uses
\(\operatorname{Map}(X,\mathbb R_q\oplus\mathbb R_p)\) with the summed
symplectic form and Stone--von Neumann uniqueness.  The fermionic test uses
CAR data on \(\operatorname{Re}(Z\oplus\overline Z)\) and the antisymmetric
Fock space.  The naive proposal that Grassmann numbers are a drop-in target
replacement is rejected in this strict Weyl-projective sense; Grassmann
variables remain a sourced GAR/fermionic phase-space calculus for later work.

Sections~\ref{sec:grassmann-weyl-fermion-displacements} and
\ref{sec:grassmann-weyl-fermion-displacement-derivations} correct the
Grassmann displacement question using Cahill--Glauber's Physical Review A
source converted with \texttt{marker}.  The abstract displacement symbols are
represented by
\(D(\gamma)=\exp(\sum_i(a_i^\dagger\gamma_i-\gamma_i^*a_i))\) inside the
Grassmann-extended finite CAR algebra.  The derivation shard proves unitarity,
the operator shifts \(a_i\mapsto a_i+\gamma_i\), and the Grassmann-valued ray
composition law, while retaining only the ordinary complex-Hilbert-space
caveat as a boundary statement.

Section~\ref{sec:prime-field-arithmetic-fields-qudit-paulis} returns to
arithmetic quantum fields in the prime-field case.  It separates
\(\operatorname{Spec}\mathbb F_p\), which has one point and gives one qudit,
from the underlying set \(\mathbb F_p\), which has \(p\) points.  For
\(X=\mathbb F_p\) as a set and \(E=\operatorname{Map}(X,\mathbb F_p^2)\), the
shard proves that the arithmetic Weyl--Heisenberg group is exactly the
standard odd-\(p\) qudit Pauli group, with the same projective labels and the
\(p=2\) fourth-root phase caveat for qubits.

Section~\ref{sec:spec-z6-residue-qudit-factorisation} then computes the first
mixed finite affine spectrum, \(R=\mathbb Z/6\mathbb Z\).  The prime ideals are
\((2)\) and \((3)\), with residue fields \(\mathbb F_2\) and \(\mathbb F_3\).
Thus the local factors are a qubit and a qutrit.  The label group is their
orthogonal direct sum, but the Weyl representation lives on
\(\mathbb C^2\otimes\mathbb C^3\), so the observable algebra is \(M_6\), not
\(M_2\oplus M_3\).

Section~\ref{sec:zariski-opens-observable-nets} adds the finite Zariski-open
observable-net convention.  The open-local assignment
\(U\mapsto\mathcal A(U)\) is proved covariant under inclusions of opens, while
the closed-complement assignment \(U\mapsto\mathcal A(X\setminus U)\) is
recorded as a useful but contravariant vanishing-locus convention.  The field
edge case has only \(\emptyset\) and \(X\); for
\(\operatorname{Spec}(\mathbb Z/6\mathbb Z)\), the singleton opens give the
commuting local subalgebras \(M_2\otimes\mathbf 1_3\) and
\(\mathbf 1_2\otimes M_3\), which generate \(M_6\).

Section~\ref{sec:quasilocal-zariski-algebra} records the directed-system
choice needed for a quasi-local algebra.  The primary construction is the
filtered colimit over quasi-compact Zariski opens,
\(\varinjlim_U\mathcal A(U)\), with inclusions following open inclusion.
The closed-complement construction is retained only after reindexing by
closed supports ordered by inclusion, as a separate vanishing-locus or defect
algebra.

Section~\ref{sec:presheaves-cosheaves-observable-nets} begins the
presheaf/cosheaf layer.  It defines presheaves, sheaves, precosheaves, and
cosheaves from local Stacks and Curry sources.  The open-local observable
assignment is proved to be a precosheaf/local net, but not an ordinary cosheaf
of unital \(*\)-algebras: for the \(\mathbb Z/6\mathbb Z\) two-point cover,
the cosheaf coproduct would be \(M_2*_{\mathbb C}M_3\), while the physical
observable algebra is \(M_2\otimes M_3\).  The Weyl label assignment
\(E(U)=\bigoplus_{x\in U}E_x\) is identified as the cosheaf-like layer.

Section~\ref{sec:state-presheaf-quasilocal-algebra} adds the state side.  The
local state assignment \(U\mapsto\operatorname{State}(\mathcal A(U))\) is a
presheaf of convex sets by restriction of functionals, equivalently by partial
trace in finite tensor products.  It is not generally a sheaf: the
\(\mathbb Z/6\mathbb Z\) two-point system gives two different global density
matrices with the same local marginals, and a three-qubit overlap example
shows that compatible marginals need not admit any global state.  For the
algebraic quasi-local algebra, global states are exactly compatible inverse
limit families of local states.

Sections~\ref{sec:closed-point-affine-arithmetic-fields},
\ref{sec:spec-fq-t-arithmetic-field}, and
\ref{sec:spec-fq-xy-arithmetic-field} extend the arithmetic-field programme
from finite spectra to affine schemes of finite type over \(\mathbb F_q\).
The new convention indexes finite Weyl/qudit sites by closed points, whose
residue fields are finite by the Nullstellensatz.  For
\(\Spec\mathbb F_q[t]\), closed points are monic irreducible polynomials and
the point \(x_\pi\) carries a \(q^{\deg\pi}\)-level qudit.  For
\(\Spec\mathbb F_q[x,y]\), rational points give a finite \(q^2\)-site
truncation, but the full quasi-local algebra ranges over all closed points of
all degrees.  Generic and curve-generic points remain geometric/topological
points rather than finite Weyl sites, and regular polynomial profiles are
recorded as formal infinite-support fields rather than quasi-local operators.

Sections~\ref{sec:weyl-heisenberg-research-status-rings},
\ref{sec:weyl-heisenberg-representations-for-kt}, and
\ref{sec:full-dual-local-models-for-kt} add the rational-function
Weyl-Heisenberg layer.  A new local source manifest registers LCA,
locally compact ring, finite Frobenius-ring, finite Clifford/Weil, local-field,
and adelic sources.  The report now separates the algebraic
\(K=k(t)\) Heisenberg group, which has an explicit irreducible unitary
Schrödinger representation on \(\ell^2(K)\), from the full LCA Stone-von
Neumann setting \(K_{\rm disc}\times\widehat{K_{\rm disc}}\) and from local or
adelic completions.  The \(k=\mathbb F_3\) examples compute concrete
\(\omega\)-commutation relations for rational-function labels.

Sections~\ref{sec:f3-polynomial-quotient-arithmetic-fields} and
\ref{sec:tn-equals-1-f3-arithmetic-field} return to concrete finite affine
examples.  For \(R_p=\mathbb F_3[t]/(p)\), the report computes the spectrum
from the irreducible factorization of \(p\), proves the Chinese-remainder and
reduced-quotient descriptions, and gives the complete residue-field
Weyl-Heisenberg representation on
\(\bigotimes_\pi\ell^2(\mathbb F_3[t]/(\pi))\).  The \(t^n=1\) shard records
the characteristic-\(3\) formula \(t^n-1=(t^m-1)^{3^{v_3(n)}}\), the
cyclotomic degree count via \(\operatorname{ord}_d(3)\), and the resulting
dimension \(\dim\mathcal H=3^m\), with nilpotent multiplicities retained as
coordinate-ring data but invisible to the present residue-qudit labels.

Sections~\ref{sec:nilpotent-thickening-weyl-fields} and
\ref{sec:thickened-tn-equals-1-examples} add the nilpotent-sensitive layer.
For a local factor \(A=\mathbb F_3[t]/(\pi^e)\), the report defines the
top-coefficient character, proves it is generating, and constructs the
projective unitary Weyl representation on \(\ell^2(A)\).  The Weyl operators
span the full matrix algebra, so the representation is irreducible.  The
\(\mathfrak m\)-adic filtration is interpreted as a jet filtration: position
jets of order \(r\) pair with momentum jets of order \(e-1-r\).  For
\(t^n=1\), this thickened layer has dimension \(3^n\), while the reduced
residue layer has dimension \(3^m\) for \(m=n/3^{v_3(n)}\).

Sections~\ref{sec:lagrangian-stabilizer-descent} and
\ref{sec:two-qutrit-stabilizer-descent-counterexample} add the stabilizer
local-to-global layer.  A phased Lagrangian label subspace is proved to give a
rank-one full stabilizer state, while vanishing stabilizer descent defect is
the separate condition that the global stabilizer labels are generated from a
chosen cover.  The two-qutrit example over \(\mathbb F_3\), generated by
\((1,0,-1,0)\) and \((0,1,0,1)\), is Lagrangian but has zero intersection
with both one-site label spaces, so its descent defect has dimension \(2\).
Thus the broad equivalence between choosing a full stabilizer and satisfying
local-to-global gluing is rejected; the corrected statement requires both a
global Lagrangian and zero descent defect, plus phase compatibility on
overlaps.

Section~\ref{sec:product-field-spectrum-qudit-stabilizers} treats the missing
finite product-field example.  For \(R=k^n\), with \(k=\mathbb F_q\), the
spectrum has \(n\) points \(\mathfrak m_i=\ker(\operatorname{pr}_i)\), each
with residue field \(k\), and the idempotents \(e_i\) make the topology
discrete.  Hence the residue-qudit field on \(X=\Spec R\) has
\[
  E(X)=k_X^n\oplus k_Z^n,\qquad
  \mathcal H(X)=\ell^2(k)^{\otimes n},\qquad
  \mathcal A(X)\cong M_{q^n}(\mathbb C).
\]
Thus \(X=\Spec(k^n)\) gives exactly the ambient \(n\) \(q\)-level qudit
Weyl/Pauli net.  Stabilizer codes remain additional data: isotropic additive
label subgroups and compatible phases, with the \(k\)-linear stabilizers as a
subclass when \(q\) is not prime.

Section~\ref{sec:cech-cohomology-quantum-descent} starts the Cech layer.  A
new Stacks Cohomology source is registered for the standard Cech complex,
\(H^0\) gluing, refinements, and ordered Cech cochains, with Curry supplying
the dual cosheaf homology convention.  The report proves that the full
residue-label cosheaf \(E(U)=\bigoplus_{x\in U}E_x\) has
\(\check H_0(\mathcal U,E)\cong E(U)\) and no higher Cech homology on finite
covers, by decomposing pointwise into full simplices.  The quantum
counterparts are then separated by layer: stabilizer descent is a genuine
abelian cokernel of zeroth Cech generation, state descent is the marginal
comparison map, and observable descent is the comparison from ordinary
algebraic cover colimits to the physical local net.

Section~\ref{sec:gaussian-clifford-dynamics} starts the dynamics layer.
Gross's arXiv TeX source for finite-dimensional Hudson theory is registered
locally.  The report fixes the odd-characteristic half-Weyl normalization,
then proves that every symplectic morphism of nondegenerate finite label
spaces induces a unital \(*\)-monomorphism of Weyl observable algebras, and
that symplectic automorphisms induce Gaussian/quasi-free automorphisms.
Gross's Clifford theorem supplies projective unitary implementations, while
Gross-Hudson identifies pure finite Gaussian states with stabilizer states.

Section~\ref{sec:finite-field-scale-continuum-limits} starts the finite-field
scale or continuum-limit layer.  For \(X_0/\mathbb F_q\), the rank-\(r\)
scale is \(X_r=X_0\times_{\mathbb F_q}\mathbb F_{q^r}\).  The report proves
that base change of schemes is canonical but naive diagonal Weyl-label
embeddings multiply absolute-trace commutators by \(s/r\).  Trace-normalized
embeddings repair this and are canonical on the prime-to-\(p\) subtower.  For
\(\Spec\mathbb F_3\), the current canonical continuum algebra is the
prime-to-\(3\) filtered colimit of \(M_{3^r}(\mathbb C)\), and Frobenius acts
by \(W(q,p)\mapsto W(q^3,p^3)\).

Section~\ref{sec:trace-gauges-full-continuum-towers} proves that the remaining
full-tower trace choices are coherent gauge data rather than an obstruction.
The key object is a trace-density system \(a_r\in\mathbb F_{q^r}\) satisfying
\(\Tr_{\mathbb F_{q^s}/\mathbb F_{q^r}}(a_s)=a_r\) for every \(r\mid s\).
Such systems exist by a factorial-tower trace-surjectivity construction, and
they define coherent normalizers \(c_{s,r}=a_s/a_r\).  Frobenius acts on this
space of gauges; for \(p\mid s/r\), no trace-one scalar fixed by the relative
Frobenius group exists.

Section~\ref{sec:spec-z-arithmetic-quantum-field} builds the first arithmetic
base-space field over \(\Spec\mathbb Z\).  The closed-prime layer has one
\(\ell\)-level qudit at each rational prime and an algebraic infinite tensor
product over finite prime supports.  The generic point is separated as a
\(\mathbb Q\)-Weyl sector rather than a finite qudit.  The report rejects
\(z\mapsto z+1\) as a scheme automorphism of \(\Spec\mathbb Z\), then defines
the internal residue-translation action \(n\mapsto\alpha_n\) on the prime-qudit
algebra, proves its profinite local periods, and shows that \(\alpha_n\) for
\(n\ne0\) is not implemented in the \(|0\rangle\) incomplete tensor sector.

\subsection*{2026-05-26 Affine Line and Regular Functions}

Sections~\ref{sec:affine-line-symmetry-weyl-net}--%
\ref{sec:f3-full-affine-line-net-hilbert-action} prove the base-preserving
symmetry group \(x\mapsto ax+b\), build its closed-point Weyl-net action, and
separate the full \(\mathbb F_3\) net from rational finite-support
truncations.  Sections~\ref{sec:regular-functions-affine-line-weyl-labels}--%
\ref{sec:f3-regular-function-symmetry-line-by-line} then identify regular
functions as residue profiles, finite equations \(h\) as support selectors,
and \(W_h^{\rm red}(F,G)\) as the actual finite Weyl smearing.  The affine
action is \(g_\#F=F\circ g^{-1}\), it transports reduced smearings by
\(\alpha_gW_h^{\rm red}(F,G)=W_{g_\#h}^{\rm red}(g_\#F,g_\#G)\), and repeated
factors remain Artin-ring thickness rather than extra closed points.
