% SHARD-ID: AQM-04-SOURCES-QUESTIONS
% SHARD-TITLE: Source Queue and Open Questions
% SHARD-SUMMARY: Lists source-acquisition priorities before substantive claims are added.
% SHARD-SUMMARY: Keeps open questions separate from checked report conclusions.
% SHARD-KEYWORDS: sources, open questions, Weil, zeta, SUSY, toric code, Levin-Wen

\section{Source Queue and Open Questions}

\subsection{Source Queue}

Registered sources now cover toric-code stabilizers, symplectic QECC
conventions, finite Hudson/Gaussian stabilizer theory, canonical CCR and CAR
field conventions, Fock spaces,
Weyl-Heisenberg and Weil representations over LCA groups, locally compact
rings, finite Frobenius rings, finite Heisenberg groups, and local/adelic
fields, and the algebraic-geometry definitions used for spectra, Zariski
topology, affine-scheme functoriality, sheaves, cosheaves, Cech cohomology,
projective space, separable trace pairing, and absolute/relative Frobenius.  Before
adding technical content in the remaining directions, register sources for:

\begin{itemize}
  \item Weil conjectures, zeta functions, Frobenius traces, and cohomological
  normalization.
  \item Arithmetic quantum mechanics and spectral interpretations of
  arithmetic data.
  \item Adelic self-dual models, Bost--Connes-style arithmetic
  \(C^*\)-dynamical systems, and any zeta-producing time evolution proposed for
  the \(\Spec\mathbb Z\) prime-qudit algebra.
  \item Supersymmetric quantum mechanics, Witten index, and Hodge/de Rham model
  conventions.
  \item Levin-Wen string-net models and categorical input data beyond the toric
  stabilizer/CSS layer.
\end{itemize}

\subsection{Open Questions}

\begin{enumerate}
  \item What is the smallest finite arithmetic example whose zeta data can be
  computed exactly and traced through the proposed dictionary?
  \item Which quantum-mechanical invariant is the first honest comparator:
  trace, index, spectrum, ground-state degeneracy, or projector rank?
  \item Which parts of the comparison are literal constructions and which parts
  are analogy or bookkeeping?
  \item What is the correct generic rational-function sector for
  \(\mathbf A^1_{\mathbb F_q}\): a discrete \(k(x)\) Weyl system with a
  dual-label affine action, a local-field completion at infinity, or an adelic
  self-dual model?
\end{enumerate}
