% SHARD-ID: AQM-10-STEANE-SYMPLECTIC-MOLECULAR
% SHARD-TITLE: Steane Code Symplectic Data in Molecular Detail
% SHARD-SUMMARY: Fixes the sourced Gottesman-table convention for the seven-qubit Steane code.
% SHARD-SUMMARY: Enumerates the binary spaces \(D\), \(C=D^\perp\), \(L\), \(L^\perp\), and \(L^\perp/L\).
% SHARD-SUMMARY: Derives the logical Pauli symplectic pairing from the explicit binary representatives.
% SHARD-KEYWORDS: Steane code, CSS code, Hamming code, symplectic reduction, stabilizer

\section{Steane Code Symplectic Data in Molecular Detail}
\label{sec:steane-symplectic-molecular}

This section fixes one concrete Steane-code presentation and writes all
finite-vector-space objects explicitly. The local source anchor is Gottesman's
\path{Thesis.tex}: lines 1249--1267 give the seven-qubit CSS stabilizer table,
and lines 1269--1284 give the two displayed encoded codewords.

\subsection{Binary Check Space}

Work over \(\mathbb F_2\) with qubits ordered \(1,\ldots,7\). The three
binary rows read from Gottesman's \(X\)-check table are
\[
\begin{aligned}
  g_1&=1111000,\\
  g_2&=1100110,\\
  g_3&=1010101.
\end{aligned}
\]
Equivalently,
\[
H=
\begin{pmatrix}
1&1&1&1&0&0&0\\
1&1&0&0&1&1&0\\
1&0&1&0&1&0&1
\end{pmatrix}.
\]
Let
\[
  D=\operatorname{row}(H)=\langle g_1,g_2,g_3\rangle\subset\mathbb F_2^7 .
\]
The eight elements of \(D\) are
\[
\begin{gathered}
0000000,\ 1111000,\ 1100110,\ 1010101,\\
0011110,\ 0101101,\ 0110011,\ 1001011 .
\end{gathered}
\]
The row dot products are
\[
HH^T=
\begin{pmatrix}
0&0&0\\
0&0&0\\
0&0&0
\end{pmatrix}
\quad \text{over } \mathbb F_2.
\]
Thus \(D\subseteq D^\perp\).

Put
\[
  C=D^\perp=\{a\in\mathbb F_2^7:Ha^T=0\}.
\]
Since \(\operatorname{rank}H=3\), \(\dim C=4\). Let
\[
  u=0000111 .
\]
Then \(u\in C\), but \(u\notin D\). Therefore
\[
  C=D\sqcup(u+D).
\]
The second coset is
\[
\begin{gathered}
0000111,\ 1111111,\ 1100001,\ 1010010,\\
0011001,\ 0101010,\ 0110100,\ 1001100 .
\end{gathered}
\]
These are exactly the two cosets supporting Gottesman's displayed
\(|\overline 0\rangle\) and \(|\overline 1\rangle\) states.

\subsection{The Stabilizer Subspace}

The Pauli label space is
\[
  V=\mathbb F_2^7\oplus\mathbb F_2^7.
\]
We write a label as \((x,z)\), corresponding to \(X^xZ^z\). The binary
symplectic form is
\[
  \omega((x,z),(x',z'))=z\cdot x'+x\cdot z',
\]
where the sign from the odd-prime formula disappears in characteristic \(2\).

The six stabilizer labels are
\[
\begin{array}{lll}
\ell_1=(g_1,0),& \ell_2=(g_2,0),& \ell_3=(g_3,0),\\
\ell_4=(0,g_1),& \ell_5=(0,g_2),& \ell_6=(0,g_3).
\end{array}
\]
Their span is
\[
  L=(D\oplus0)+(0\oplus D)\subset V.
\]
This is an isotropic subspace. Indeed, same-type pairings vanish automatically.
For a mixed pair,
\[
  \omega((d,0),(0,d'))=d\cdot d'=0
  \qquad(d,d'\in D),
\]
because \(D\subseteq D^\perp\). Therefore \(L\subseteq L^\perp\).

The dimensions and sizes are
\[
  \dim L=6,\qquad |L|=2^6=64.
\]

\subsection{The Centralizer}

Let \((x,z)\in V\). It is orthogonal to every \((d,0)\in D\oplus0\) exactly
when
\[
  z\cdot d=0\quad\text{for all }d\in D,
\]
that is, \(z\in C\). It is orthogonal to every \((0,d)\in0\oplus D\) exactly
when
\[
  x\cdot d=0\quad\text{for all }d\in D,
\]
that is, \(x\in C\). Hence
\[
  L^\perp=C\oplus C.
\]
Consequently
\[
  \dim L^\perp=8,\qquad |L^\perp|=2^8=256.
\]

\subsection{The Logical Symplectic Quotient}

The logical Pauli label space is the finite symplectic reduction
\[
  L^\perp/L
  =
  (C\oplus C)/(D\oplus D)
  \cong
  (C/D)\oplus(C/D).
\]
Since \(C=D\sqcup(u+D)\), the quotient \(C/D\) is one-dimensional. Thus
\[
  \dim(L^\perp/L)=2.
\]
Choose the representatives
\[
  \overline X=(u,0),\qquad \overline Z=(0,u).
\]
Their symplectic pairing is
\[
  \omega(\overline X,\overline Z)
  =
  u\cdot u
  =
  1+1+1
  =
  1
  \quad \text{in }\mathbb F_2.
\]
Thus the corresponding Pauli operators anticommute. The four logical Pauli
classes are represented by
\[
  (0,0),\quad (u,0),\quad (0,u),\quad (u,u).
\]

This is the full symplectic anatomy of the code: \(D\) supplies the CSS check
space, \(L=D\oplus D\) is the isotropic stabilizer subspace, \(L^\perp=C\oplus
C\) is the Pauli centralizer, and \(L^\perp/L\) is the one-qubit logical
symplectic module.
