% SHARD-ID: AQM-11-STEANE-SUPERCHARGE-COHOMOLOGY
% SHARD-TITLE: Steane Code Fermions, Supercharge, and Cohomology
% SHARD-SUMMARY: Constructs the six-fermion Koszul supercharge for the sourced Steane stabilizer generators.
% SHARD-SUMMARY: Proves directly that the no-ghost cohomology is the two-dimensional Steane code.
% SHARD-SUMMARY: Computes the full ghost-degree cohomology and separates it from the physical code space.
% SHARD-KEYWORDS: Steane code, supercharge, fermions, Koszul complex, cohomology, stabilizer Hamiltonian

\section{Steane Code Fermions, Supercharge, and Cohomology}
\label{sec:steane-supercharge-cohomology}

Keep the notation of Section~\ref{sec:steane-symplectic-molecular}. This
section constructs the actual fermions and supercharge for the same six
Gottesman-table stabilizer generators.

\subsection{Physical Stabilizer Projectors}

The physical Hilbert space is
\[
  \mathcal H=\mathbb C\{|a\rangle:a\in\mathbb F_2^7\}.
\]
For \(d\in\mathbb F_2^7\), define
\[
  X(d)|a\rangle=|a+d\rangle,\qquad
  Z(d)|a\rangle=(-1)^{d\cdot a}|a\rangle .
\]
The six stabilizer operators are
\[
\begin{array}{lll}
S_1=X(g_1),&S_2=X(g_2),&S_3=X(g_3),\\
S_4=Z(g_1),&S_5=Z(g_2),&S_6=Z(g_3).
\end{array}
\]
The violated-check projectors are
\[
  P_i=\frac{I-S_i}{2},\qquad
  \Pi_i=\frac{I+S_i}{2}.
\]
Since the labels span the isotropic subspace \(L\), all \(S_i\), hence all
\(P_i\), commute.

\subsection{The Two Codeword Cosets}

The two normalized Steane basis states in this convention are
\[
  |\overline0\rangle
  =
  \frac1{\sqrt8}\sum_{d\in D}|d\rangle,
  \qquad
  |\overline1\rangle
  =
  \frac1{\sqrt8}\sum_{d\in D}|u+d\rangle .
\]
For an \(X\)-type check \(X(g_j)\), adding \(g_j\in D\) permutes \(D\) and also
permutes \(u+D\). Hence
\[
  X(g_j)|\overline0\rangle=|\overline0\rangle,\qquad
  X(g_j)|\overline1\rangle=|\overline1\rangle .
\]
For a \(Z\)-type check \(Z(g_j)\), every vector in \(D\) is orthogonal to
\(g_j\), and \(u\) is also orthogonal to \(g_j\). Therefore
\[
  Z(g_j)|d\rangle=|d\rangle,\qquad
  Z(g_j)|u+d\rangle=|u+d\rangle .
\]
So both displayed states are fixed by all six stabilizers.

Conversely, let
\[
  |\psi\rangle=\sum_{a\in\mathbb F_2^7}\psi_a|a\rangle .
\]
The equations \(Z(g_j)|\psi\rangle=|\psi\rangle\) force \(\psi_a=0\) unless
\(a\in C=D^\perp\). The equations \(X(g_j)|\psi\rangle=|\psi\rangle\) force
\(\psi_{a+g_j}=\psi_a\), so the amplitudes are constant on \(D\)-cosets inside
\(C\). There are exactly two such cosets, \(D\) and \(u+D\). Thus the common
stabilizer \(+1\) space is
\[
  \mathcal C=\operatorname{span}\{|\overline0\rangle,|\overline1\rangle\},
  \qquad \dim\mathcal C=2.
\]

\subsection{Six Fermions}

Let
\[
  E=\operatorname{span}_{\mathbb C}
  \{\epsilon_1,\ldots,\epsilon_6\}.
\]
The ghost Fock space is the exterior algebra \(\Lambda E\). Exterior
multiplication by \(\epsilon_i\) and contraction \(\iota_i\) obey
\[
  \iota_i\epsilon_j+\epsilon_j\iota_i=\delta_{ij},
  \qquad
  \epsilon_i\epsilon_j+\epsilon_j\epsilon_i=0,
  \qquad
  \iota_i\iota_j+\iota_j\iota_i=0.
\]
The ghost-extended Hilbert space is
\[
  \mathcal K=\mathcal H\otimes\Lambda E,
  \qquad
  \dim\mathcal K=2^7\,2^6=8192.
\]

Define
\[
  Q=\sum_{i=1}^6\epsilon_iP_i,\qquad
  Q^*=\sum_{i=1}^6\iota_iP_i .
\]
Written without abbreviations,
\[
\begin{aligned}
Q={}&
\epsilon_1\frac{I-X(g_1)}2+
\epsilon_2\frac{I-X(g_2)}2+
\epsilon_3\frac{I-X(g_3)}2\\
&+
\epsilon_4\frac{I-Z(g_1)}2+
\epsilon_5\frac{I-Z(g_2)}2+
\epsilon_6\frac{I-Z(g_3)}2 .
\end{aligned}
\]
Because the \(P_i\) commute and the \(\epsilon_i\) anticommute,
\[
  Q^2=\sum_{i<j}(\epsilon_i\epsilon_j+\epsilon_j\epsilon_i)P_iP_j=0.
\]
The supersymmetric Hamiltonian is
\[
\begin{aligned}
  QQ^*+Q^*Q
  &=\sum_{i,j}(\epsilon_i\iota_j+\iota_j\epsilon_i)P_iP_j\\
  &=\sum_iP_i .
\end{aligned}
\]
Thus it is the Steane stabilizer-violation Hamiltonian, shifted and scaled so
that the code sector has energy \(0\).

\subsection{Cohomology, Defined and Computed}

Ghost degree is exterior degree. The degree-\(q\) subspace is
\[
  \mathcal K^q=\mathcal H\otimes\Lambda^qE.
\]
The degree-\(q\) cohomology of \(Q\) is
\[
  H^q(Q)=
  \frac{\{\phi\in\mathcal K^q:Q\phi=0\}}
       {\{Q\eta:\eta\in\mathcal K^{q-1}\}} .
\]
For \(q=0\), the denominator is zero because \(\mathcal K^{-1}=0\).

Decompose the physical Hilbert space into stabilizer-syndrome sectors
\[
  \mathcal H=\bigoplus_{\lambda\in\mathbb F_2^6}\mathcal H_\lambda,
  \qquad
  S_i\psi=(-1)^{\lambda_i}\psi .
\]
On \(\mathcal H_\lambda\), the projector \(P_i\) acts as multiplication by
\(\lambda_i\). Therefore
\[
  Q|_{\mathcal H_\lambda\otimes\Lambda E}
  =
  \epsilon(\alpha_\lambda),
  \qquad
  \alpha_\lambda=\sum_i\lambda_i\epsilon_i,
\]
where \(\epsilon(\alpha)\) means exterior multiplication by \(\alpha\).

If \(\lambda\neq0\), choose a contraction \(h\) with \(h(\alpha_\lambda)=1\).
Then
\[
  hQ+Qh=I
\]
on \(\mathcal H_\lambda\otimes\Lambda E\), so that syndrome block is acyclic:
it contributes no cohomology. If \(\lambda=0\), then \(Q=0\), and
\(\mathcal H_0=\mathcal C\) has dimension \(2\). Hence
\[
  H^q(Q)\cong\mathcal C\otimes\Lambda^qE,
  \qquad
  \dim H^q(Q)=2\binom6q .
\]
The dimensions are
\[
\begin{array}{c|ccccccc}
q&0&1&2&3&4&5&6\\
\hline
\dim H^q(Q)&2&12&30&40&30&12&2 .
\end{array}
\]

The physical Steane code is the no-ghost part:
\[
  H^0(Q)=\mathcal C
  =
  \operatorname{span}\{|\overline0\rangle,|\overline1\rangle\}.
\]
The full \(Q\)-cohomology has ghost-labelled copies of the same code space,
with total dimension \(2\cdot2^6=128\). The ground space of
\(\{Q,Q^*\}\) has the same ghost degeneracy. The no-ghost sector is therefore
the physical stabilizer code; the higher ghost-degree classes are auxiliary
Koszul bookkeeping.

\subsection{What the Higher Ghost Classes Mean}

The word ``bookkeeping'' in the previous paragraph should be read narrowly. It
means that higher ghost degree does not add new Steane logical qubits. It does
not mean that the Koszul cohomology is meaningless.

The six equations defining the code are
\[
  P_1\psi=\cdots=P_6\psi=0.
\]
The Koszul differential remembers these as six separate equations. On a
nonzero syndrome sector \(\lambda\neq0\), at least one equation is violated and
the complex is contractible by the homotopy \(h\) above. On the zero-syndrome
sector, all six functions \(P_i\) vanish. Therefore \(Q\) becomes the zero
operator on
\[
  \mathcal C\otimes\Lambda E,
\]
and every exterior ghost monomial survives.

Thus the higher cohomology is exactly
\[
  H^{>0}(Q)=\mathcal C\otimes\Lambda^{>0}E.
\]
It records the exterior algebra on the six chosen stabilizer equations over the
constraint surface. Equivalently, it is the finite-dimensional analogue of
conormal or derived-intersection data for the equations \(P_i=0\). This data is
useful if the question is about the Koszul resolution itself, or about how the
constraint presentation changes when generators are added, removed, or made
redundant.

It is not, however, an invariant of the Steane code alone. If we choose a
different independent generating set for the same \(L\), the ghost basis
changes. If we add redundant stabilizer generators, the higher exterior algebra
gets larger while the physical code space is unchanged. The intrinsic QECC
content is therefore the no-ghost code \(\mathcal C=H^0(Q)\) together with the
logical symplectic quotient \(L^\perp/L\). The full \(Q\)-cohomology is a
presentation-dependent derived enhancement of the stabilizer equations.

\subsection{Exact Validation}

The producer \path{scripts/lattice_codes/steane_supercharge_molecular.jl}
writes the CSV files
\path{runs/2026-05-24-steane-supercharge-molecular/data/steane_molecular_summary.csv},
\path{runs/2026-05-24-steane-supercharge-molecular/data/steane_molecular_vectors.csv},
and
\path{runs/2026-05-24-steane-supercharge-molecular/data/steane_cohomology_by_degree.csv}.
The summary CSV records \(\dim D=3\), \(\dim C=4\), \(\dim L=6\),
\(\dim L^\perp=8\), \(|\mathcal K|=8192\), \(\dim H^0(Q)=2\), and full
cohomology dimension \(128\). The vector CSV records the two codeword cosets
and the logical representatives \((u,0)\), \((0,u)\). The cohomology CSV records
the seven dimensions displayed above.
