% SHARD-ID: AQM-12-STEANE-CLIFFORD-KOSZUL-MORPHISMS
% SHARD-TITLE: Steane Clifford Morphisms of the Koszul Supercharge
% SHARD-SUMMARY: Defines the Clifford lift of a symplectic Pauli-label automorphism to the Steane Koszul complex.
% SHARD-SUMMARY: Proves that transversal Hadamard gives an exact chain automorphism after ghost swapping.
% SHARD-SUMMARY: Shows that transversal phase gives a chain isomorphism to a changed presentation and a homotopy equivalence back.
% SHARD-KEYWORDS: Steane code, Clifford, symplectic automorphism, Koszul complex, homotopy equivalence

\section{Steane Clifford Morphisms of the Koszul Supercharge}
\label{sec:steane-clifford-koszul-morphisms}

Here ``symplectic morphism'' means a Clifford operation: a unitary \(U\) whose
conjugation action on Paulis induces a symplectic automorphism \(F_U\) of the
binary label space. Gottesman's local source anchors are \path{Thesis.tex}:
lines 2271--2305 for valid encoded Clifford operations preserving the
stabilizer, lines 2307--2348 for the Hadamard, phase, and CNOT Pauli-label
actions, and lines 2439--2445 for bitwise Hadamard and phase on the
seven-qubit code.

\subsection{General Clifford Lift}

Let \(G=(\ell_1,\ldots,\ell_6)\) be the Steane generator list from
Section~\ref{sec:steane-symplectic-molecular}, and let
\[
  Q_G=\sum_i\epsilon_iP_{\ell_i},
  \qquad
  P_{\ell_i}=\frac{I-W(\ell_i)}2 .
\]
If a Clifford \(U\) sends \(W(\ell)\) to \(W(F_U\ell)\), up to the stabilizer
phase convention, then
\[
  U P_{\ell_i} U^\dagger=P_{F_U\ell_i}.
\]
Let \(G'=F_U(G)\), give \(G'\) ghosts \(\epsilon'_i\), and define
\[
  Q_{G'}=\sum_i\epsilon'_iP_{F_U\ell_i}.
\]
The map
\[
  \Phi_U=U\otimes\Lambda T,\qquad T(\epsilon_i)=\epsilon'_i,
\]
satisfies
\[
  \Phi_U Q_G=Q_{G'}\Phi_U.
\]
This is a chain isomorphism, hence it induces isomorphisms
\[
  H^q(Q_G)\cong H^q(Q_{G'}).
\]
If \(G'\) is just a permutation of \(G\), and \(T\) is the same permutation of
ghosts, this is an automorphism of the same written operator \(Q_G\). If \(G'\)
is a different independent basis of the same \(L\), it is instead an
isomorphism to a different presentation of the same stabilizer constraints.

\subsection{Transversal Hadamard}

For bitwise Hadamard \(R^{\otimes7}\),
\[
  F_H(x,z)=(z,x).
\]
Therefore
\[
  (g_i,0)\mapsto(0,g_i),\qquad
  (0,g_i)\mapsto(g_i,0).
\]
So
\[
  \ell_1\leftrightarrow\ell_4,\quad
  \ell_2\leftrightarrow\ell_5,\quad
  \ell_3\leftrightarrow\ell_6.
\]
With the ghost swap
\[
  \epsilon_1\leftrightarrow\epsilon_4,\quad
  \epsilon_2\leftrightarrow\epsilon_5,\quad
  \epsilon_3\leftrightarrow\epsilon_6,
\]
we get the exact chain automorphism
\[
  (R^{\otimes7}\otimes\Lambda T_H)Q_G
  =
  Q_G(R^{\otimes7}\otimes\Lambda T_H).
\]
On the logical symplectic quotient this sends
\[
  \overline X=(u,0)\mapsto(0,u)=\overline Z,
  \qquad
  \overline Z=(0,u)\mapsto(u,0)=\overline X.
\]
On the no-ghost code basis,
\[
  R^{\otimes7}|\overline0\rangle
  =
  \frac{|\overline0\rangle+|\overline1\rangle}{\sqrt2},
  \qquad
  R^{\otimes7}|\overline1\rangle
  =
  \frac{|\overline0\rangle-|\overline1\rangle}{\sqrt2}.
\]
Thus the induced map on \(H^0(Q_G)\) is the encoded Hadamard, and on
\(H^q(Q_G)\cong\mathcal C\otimes\Lambda^qE\) it is encoded Hadamard tensored
with the ghost swap.

\subsection{Transversal Phase}

For bitwise phase \(P^{\otimes7}\), ignoring Pauli phases in the binary label,
\[
  F_P(x,z)=(x,z+x).
\]
Hence
\[
  (g_i,0)\mapsto(g_i,g_i)=\ell_i+\ell_{i+3},
  \qquad
  (0,g_i)\mapsto(0,g_i)=\ell_{i+3}.
\]
These six images are an independent basis of the same \(L\), but they are not
the original generator list and not a permutation of it. Therefore
\[
  P^{\otimes7}\otimes\Lambda T_P
\]
is a chain isomorphism
\[
  (\mathcal H\otimes\Lambda E,Q_G)
  \cong
  (\mathcal H\otimes\Lambda E',Q_{F_P(G)}),
\]
not literally an automorphism of the original written \(Q_G\).

To compare \(Q_{F_P(G)}\) back with \(Q_G\), decompose into the same physical
syndrome sectors. The zero-syndrome sector is the Steane code. Every nonzero
syndrome sector is contractible, as in Section~\ref{sec:steane-supercharge-cohomology}.
Both complexes therefore deformation retract to their zero-syndrome
cohomology model
\[
  \mathcal C\otimes\Lambda E
\]
after choosing the corresponding ghost basis. This is the precise homotopy
equivalence: it is presentation-dependent on ghosts, but identical on the
physical code sector. On logical labels,
\[
  \overline X=(u,0)\mapsto(u,u)=\overline X+\overline Z,
  \qquad
  \overline Z=(0,u)\mapsto(0,u)=\overline Z.
\]

\subsection{Conclusion}

For Steane, Clifford symmetry gives more than an abstract equivalence relation.
It gives a chain isomorphism from the Koszul complex of one stabilizer
presentation to the Koszul complex of the transformed presentation. When the
Clifford permutes the chosen generators, as transversal Hadamard does here,
this is an exact automorphism of \(Q_G\). When the Clifford changes the
independent generator basis, as transversal phase does here, the complexes are
homotopy equivalent through their common zero-syndrome deformation retract, but
the full higher-ghost cohomology remains presentation-labelled.

\subsection{Run Artifact Record}

The producer
\path{scripts/lattice_codes/steane_clifford_koszul_morphisms.jl} writes the
Steane Clifford/Koszul run bundle
\path{runs/2026-05-24-steane-clifford-koszul-morphisms/}. Its summary CSV
records Boolean and string fields for the Hadamard generator permutation, the
phase generator-basis change, the binary logical-label images, the count of
\(63\) nonzero syndrome sectors, the summary contractibility flag, and the
cohomology dimensions listed in
Section~\ref{sec:steane-supercharge-cohomology}. The companion generator-map
CSV lists the six stabilizer-label images and their coordinates in the original
generator basis. These artifacts record checked hypotheses and summary
consequences used by the local derivation above; they do not contain full
chain-map matrices, contraction homotopies, or residual identities.
