% SHARD-ID: AQM-13-CLIFFORD-GROUP-GHOST-GAUSSIAN-THEOREM
% SHARD-TITLE: Clifford Group Ghost-Gaussian Theorem
% SHARD-SUMMARY: States and proves the arbitrary-Clifford covariance theorem for signed Pauli Koszul supercharges.
% SHARD-SUMMARY: Proves that stabilizer presentation changes lift to number-preserving fermionic Gaussian ghost operations.
% SHARD-SUMMARY: Applies the theorem to all standard Clifford generators on the Steane stabilizer presentation.
% SHARD-KEYWORDS: Clifford group, Pauli group, signed stabilizers, ghost Gaussian, Koszul complex, Steane code

\section{Clifford Group Ghost-Gaussian Theorem}
\label{sec:clifford-group-ghost-gaussian-theorem}

The previous section treated two transversal Steane Cliffords.  The correct
general statement is stronger, but it needs one extra convention: binary Pauli
labels alone are not enough.  Gottesman's local source says that Clifford
normalizer elements conjugate Pauli products to Pauli products
\path{Thesis.tex}, lines 2292--2305 and 2307--2348; the binary stabilizer
language drops phases at lines 1303--1327.  The source manifest records the
full local path.  The supercharge projectors do not allow that drop.

\subsection{Signed Pauli Convention}

For \(n\) qubits define the Hermitian representative
\[
  W(x,z)=i^{x\cdot z}\bigotimes_{j=1}^n X_j^{x_j}Z_j^{z_j},
  \qquad x,z\in\mathbb F_2^n .
\]
A signed Pauli stabilizer generator is
\[
  S=\tau W(x,z),\qquad \tau\in\{\pm1\}.
\]
The violated-check projector is
\[
  P_S=\frac{I-S}{2}.
\]
Thus if a Clifford gives
\[
  U W(x,z)U^\dagger=\tau W(x',z'),
\]
then the transported projector is
\[
  U P_{W(x,z)}U^\dagger=\frac{I-\tau W(x',z')}{2}.
\]
When \(\tau=-1\), this is the opposite syndrome projector for the unsigned
binary label \((x',z')\).  This is why the theorem below is stated for signed
Pauli operators, not only for symplectic binary labels.

\subsection{The Koszul Complex Attached to a Signed Presentation}

Let
\[
  G=(S_1,\ldots,S_r)
\]
be an ordered independent list of commuting signed Hermitian Pauli operators,
with no product equal to \(-I\).  Let \(E=\mathrm{span}_{\mathbb C}
\{\epsilon_1,\ldots,\epsilon_r\}\), and let \(c_i^\dagger\) denote exterior
multiplication by \(\epsilon_i\) on \(\Lambda E\).  The ghost-extended complex
is
\[
  \mathcal K(G)=\mathcal H\otimes\Lambda E,
  \qquad
  Q_G=\sum_{i=1}^r c_i^\dagger P_{S_i}.
\]
This is exactly the earlier supercharge notation
\(\sum_i\epsilon_iP_{S_i}\), now with the creation operator written
explicitly.  Since the \(P_{S_i}\) commute and the \(c_i^\dagger\) anticommute,
\(Q_G^2=0\).

\subsection{Theorem: Arbitrary Clifford Covariance}

Let \(U\) be any \(n\)-qubit Clifford.  Define the transported signed
presentation
\[
  U(G)=\bigl(US_1U^\dagger,\ldots,US_rU^\dagger\bigr)
\]
and give it ghosts \(E^U=\mathrm{span}\{\epsilon_1^U,\ldots,\epsilon_r^U\}\).
Let \(T_U:E\to E^U\) be the linear map \(T_U(\epsilon_i)=\epsilon_i^U\).  Then
\[
  \Phi_U=U\otimes\Lambda T_U
\]
is a chain isomorphism
\[
  \Phi_U Q_G=Q_{U(G)}\Phi_U .
\]
Equivalently,
\[
  Q_{U(G)}
  =
  (U\otimes\Lambda T_U)Q_G(U^\dagger\otimes\Lambda T_U^{-1}).
\]

\emph{Proof.}
For each \(i\),
\[
  U P_{S_i} U^\dagger
  =
  U\frac{I-S_i}{2}U^\dagger
  =
  \frac{I-US_iU^\dagger}{2}
  =
  P_{US_iU^\dagger}.
\]
Now apply both sides to a pure tensor \(\psi\otimes\omega\):
\[
\begin{aligned}
  \Phi_UQ_G(\psi\otimes\omega)
  &=
  \sum_i U(P_{S_i}\psi)\otimes
        \epsilon_i^U\wedge(\Lambda T_U)\omega                 \\
  &=
  \sum_i P_{US_iU^\dagger}(U\psi)\otimes
        \epsilon_i^U\wedge(\Lambda T_U)\omega                 \\
  &=
  Q_{U(G)}\Phi_U(\psi\otimes\omega).
\end{aligned}
\]
The map is invertible because \(U\) and \(T_U\) are invertible.  This proves
the chain isomorphism and hence gives canonical isomorphisms
\[
  H^q(Q_G)\cong H^q(Q_{U(G)})
\]
for every ghost degree \(q\).

\subsection{Where the Ghost Gaussian Enters}

The covariance theorem uses the transported ordered list \(U(G)\), so the ghost
map is just \(\epsilon_i\mapsto\epsilon_i^U\).  The genuinely nontrivial ghost
operation appears when we rewrite a signed stabilizer group in a different
independent generator basis.

First consider a row swap: exchange \(S_a\) and \(S_b\).  The supercharge is
unchanged after the same swap of ghost creation operators,
\[
  c_a^\dagger\leftrightarrow c_b^\dagger .
\]
This is a number-preserving fermionic Gaussian operation on the ghost Fock
space.

Now consider the elementary row shear
\[
  S_a\longmapsto S_aS_b,\qquad S_i\longmapsto S_i\quad(i\ne a).
\]
The key projector identity is not linear over \(\mathbb C\) in the syndrome
bits; it contains the old stabilizer operator:
\[
\begin{aligned}
  P_{S_aS_b}
  &=
  \frac{I-S_aS_b}{2}                                      \\
  &=
  \frac{I-S_a}{2}+S_a\frac{I-S_b}{2}                       \\
  &=
  P_{S_a}+S_aP_{S_b}.
\end{aligned}
\]
Therefore the sheared supercharge is
\[
\begin{aligned}
  Q'
  &=
  c_a^\dagger P_{S_aS_b}
  +c_b^\dagger P_{S_b}
  +\sum_{i\ne a,b}c_i^\dagger P_{S_i}                       \\
  &=
  c_a^\dagger P_{S_a}
  +(c_b^\dagger+S_ac_a^\dagger)P_{S_b}
  +\sum_{i\ne a,b}c_i^\dagger P_{S_i}.
\end{aligned}
\]
Thus the ghost creation operators transform by
\[
  c_b^\dagger\longmapsto c_b^\dagger+S_ac_a^\dagger,
  \qquad c_i^\dagger\longmapsto c_i^\dagger\quad(i\ne b).
\]
This is a number-preserving fermionic Gaussian with an even
operator-valued coefficient:
\[
  \Gamma_{ab}=\exp(S_a c_a^\dagger c_b).
\]
Indeed \(S_a\) commutes with all projectors and ghost operators, and the
commutator gives
\[
  \Gamma_{ab}c_b^\dagger\Gamma_{ab}^{-1}
  =
  c_b^\dagger+S_ac_a^\dagger,
  \qquad
  \Gamma_{ab}c_i^\dagger\Gamma_{ab}^{-1}=c_i^\dagger
  \quad(i\ne b).
\]
Consequently
\[
  Q'=\Gamma_{ab}Q_G\Gamma_{ab}^{-1}.
\]

Row swaps and row shears generate \(\mathrm{GL}_r(\mathbb F_2)\).  Therefore
every independent presentation
\[
  S'_a=\prod_i S_i^{A_{ai}},\qquad A\in\mathrm{GL}_r(\mathbb F_2),
\]
is obtained by composing these elementary ghost Gaussians.  The closed
projector formula for a row \(I=(i_1,\ldots,i_m)\) is the telescoping identity
\[
  P_{\prod_{\nu=1}^mS_{i_\nu}}
  =
  \sum_{\mu=1}^m
  \left(\prod_{\nu<\mu}S_{i_\nu}\right)P_{S_{i_\mu}} .
\]
This formula is the concrete reason that a scalar exterior
\(\mathrm{GL}_r(\mathbb F_2)\) operation on ghosts is not enough on the full
physical Hilbert space.  The syndrome parity
\(\sigma_a\oplus\sigma_b\) is represented over \(\mathbb C\) by
\[
  \sigma_a+(-1)^{\sigma_a}\sigma_b,
\]
not by \(\sigma_a+\sigma_b\).  The coefficient \((-1)^{\sigma_a}\) is exactly
the eigenvalue of \(S_a\).

On the code subspace all \(S_i=+1\).  Hence every row shear reduces there to
the ordinary exterior linear map
\[
  c_b^\dagger\mapsto c_b^\dagger+c_a^\dagger .
\]
So the higher-ghost cohomology is functorial under the exterior representation
of the chosen presentation change on the zero-syndrome sector, while the full
chain-level comparison remembers operator-valued syndrome signs.

\subsection{Steane: All Clifford Generators}

For the Steane code, \(r=6\).  The standard Clifford generating gates are the
seven single-qubit \(H\)'s, seven single-qubit \(P\)'s, and the \(42\) ordered
CNOT gates.  This is the generating set described in the local Gottesman source
at lines 2292--2348.  The producer
\path{scripts/lattice_codes/steane_all_clifford_ghost_gaussians.jl} applies
these \(56\) gates to the six sourced Steane stabilizer generators from
Section~\ref{sec:steane-symplectic-molecular}.

The run records \(336\) signed generator images.  Every one of the \(56\)
image lists has binary rank \(6\) and is isotropic for the binary symplectic
form.  Since the listed \(H\), \(P\), and CNOT gates generate the Clifford
group, any Clifford word is handled by composing these signed Pauli-label
updates and then applying the covariance theorem above.

The same run records the \(45\) elementary ghost-Gaussian presentation moves
for the Steane six-generator group: \(15\) row swaps and \(30\) directed row
shears.  For each shear it checks all \(64\) Steane syndrome sectors, i.e.
\[
  30\cdot64=1920
\]
instances of
\[
  P(S_aS_b)=P(S_a)+S_aP(S_b).
\]

If a Clifford \(U\) lies in the Steane stabilizer normalizer, then
\[
  US_iU^\dagger=\prod_jS_j^{A_{ij}}
\]
for some \(A\in\mathrm{GL}_6(\mathbb F_2)\).  In that case
\[
  (U\otimes I)Q_G(U^\dagger\otimes I)=Q_{A(G)}
  =
  \Gamma_AQ_G\Gamma_A^{-1},
\]
where \(\Gamma_A\) is any product of the elementary ghost Gaussians above that
realizes the row-reduction path for \(A\).  If \(U\) is an arbitrary Clifford
not preserving the Steane stabilizer group, the first equality still gives a
strict supercharge for the transported code \(U\mathcal C\), but no ghost-only
operation can turn its physical projectors back into the original Steane
projectors.

\subsection{Conclusion}

The precise answer is therefore Clifford conjugation on physical qubits plus a
fermionic Gaussian change of ghost presentation.  No Bogoliubov mixing of
creation and annihilation operators is needed for the cohomological complex;
the transformations preserve ghost degree.  The only non-scalar feature is
essential: when a product stabilizer is rewritten in terms of old violated-check
projectors, the Gaussian coefficients are even operators generated by the old
stabilizers themselves.
