% SHARD-ID: AQM-25-QUASILOCAL-ZARISKI-ALGEBRA
% SHARD-TITLE: Quasi-Local Zariski Algebra
% SHARD-SUMMARY: Restricts the AQM-24 quasi-local construction to finite spectra and finite supports.
% SHARD-SUMMARY: Records that quasi-compact opens in affine spectra have a terminal global open and therefore are not the infinite affine finite-region spine.
% SHARD-SUMMARY: Points infinite affine locality to the AQM-28 finite closed-support construction.
% SHARD-KEYWORDS: quasi-local algebra, directed system, filtered colimit, finite spectrum, finite support, closed support

\section{Quasi-Local Zariski Algebra}
\label{sec:quasilocal-zariski-algebra}

\subsection{Status and Source Anchors}

This shard records the finite-spectrum answer to the directed-system question
left by AQM-24.  It is restricted to the same residue-qudit setting as AQM-24:
finite spectra with finite local Weyl factors and tensor-by-identity
inclusions.  It does not define a general infinite affine quasi-local algebra
by taking a colimit over quasi-compact opens.

The Zariski-topology source remains
\path{references/algebraic_geometry/stacks_project_algebra.tex}: standard
opens form a basis in lines 3104--3117, \(\operatorname{Spec}R\) is
quasi-compact in lines 3251--3273, and the topology has a quasi-compact open
basis with quasi-compact intersections in lines 3275--3298.

No analytic completion convention is fixed here.  The object constructed below
is the algebraic finite-support \(*\)-algebra.  Because \(X\) is finite in
this shard, that algebra has the global finite tensor product \(\mathcal A(X)\)
as a terminal value.  The later infinite affine finite-region construction is
AQM-28: finite closed supports inside \(U\cap |X|_{\rm cl}\).

\subsection{Lamport Dependency Skeleton}

\[
\begin{array}{rcl}
D1&:&\text{A directed set is a preorder in which any two elements have
      a common upper bound.}\\
D2&:&X\text{ is a finite residue-qudit spectrum as in AQM-24}.\\
D3&:&\mathsf F(X)=\{S\subset X:S\text{ finite}\},\quad
      \mathcal A(S)=\bigotimes_{x\in S}\mathcal A_x.\\
L1&:&\mathsf F(X)\text{ is directed under inclusion.}\\
L2&:&S\subset T\Rightarrow
      \mathcal A(S)\hookrightarrow\mathcal A(T)
      \text{ functorially by tensoring identities.}\\
T1&:&\varinjlim_{S\in\mathsf F(X)}\mathcal A(S)=\mathcal A(X)
      \text{ for finite }X.\\
T2&:&\text{Quasi-compact opens in affine spectra have a global terminal
      open and are not the infinite finite-region index.}\\
T3&:&\text{Infinite affine locality is the finite closed-support system of
      AQM-28.}
\end{array}
\]

\subsection{Finite-Support Poset D1--L2}

A preorder \(I\) is called directed if for every \(i,j\in I\) there exists
\(k\in I\) such that \(i\leq k\) and \(j\leq k\).  A directed system of
\(*\)-algebras over \(I\) is a family \((A_i)_{i\in I}\) with structure maps
\(\alpha_{ij}:A_i\to A_j\) for \(i\leq j\), satisfying
\[
  \alpha_{ii}=\operatorname{id},\qquad
  \alpha_{jk}\alpha_{ij}=\alpha_{ik}.
\]

Let \(X=\operatorname{Spec}R\) be one of the finite residue-qudit spectra of
AQM-24.  For a finite support \(S\subset X\), AQM-24 defines
\[
  \mathcal A(S)=\bigotimes_{x\in S}\mathcal A_x
\]
with empty tensor product \(\mathbb C\).  Write
\[
  \mathsf F(X)=\{S\subset X:S\text{ finite}\}
\]
ordered by inclusion.  Since \(X\) is finite here, every subset is finite.

\paragraph{Lemma \(L1\).}
\(\mathsf F(X)\) is directed under inclusion.

\emph{Proof.}
If \(S,T\subset X\) are finite, then \(S\cup T\) is finite and contains both.
Thus \(S\cup T\) is a common upper bound.

\paragraph{Lemma \(L2\).}
The finite-support algebras form a directed system.

\emph{Proof.}
For \(S\subset T\), use AQM-24's inclusion
\[
  \iota_{S,T}:\mathcal A(S)\hookrightarrow\mathcal A(T),\qquad
  a\mapsto a\otimes\mathbf 1_{T\setminus S}.
\]
AQM-24 proves that these maps are injective unital \(*\)-homomorphisms and
that \(\iota_{T,U}\iota_{S,T}=\iota_{S,U}\) for \(S\subset T\subset U\).
Together with Lemma \(L1\), these are exactly the directed-system axioms.

\subsection{Finite-Spectrum Quasi-Local System T1}

Define the finite-spectrum algebraic quasi-local algebra by
\[
  \mathcal A_{\rm finqloc}(X)
  :=
  \varinjlim_{S\in\mathsf F(X)}\mathcal A(S).
\]
For an open \(U\subset X\), use the same construction inside \(U\):
\[
  \mathcal A_{\rm finqloc}(U)
  :=
  \varinjlim_{S\in\mathsf F(U)}\mathcal A(S).
\]
Since \(X\) and \(U\) are finite in this shard, \(X\) and \(U\) are terminal
objects in their finite-support posets.

\paragraph{Theorem \(T1\).}
\[
  \mathcal A_{\rm finqloc}(X)\cong\mathcal A(X),\qquad
  \mathcal A_{\rm finqloc}(U)\cong\mathcal A(U).
\]
Thus AQM-25 supplies the finite-spectrum analogue of a quasi-local algebra,
not an infinite affine finite-region construction.

\emph{Proof.}
The colimit over a directed system with terminal object is the terminal value.
Here \(X\) is terminal in \(\mathsf F(X)\), and \(U\) is terminal in
\(\mathsf F(U)\).  The comparison maps are the AQM-24 inclusions into the
terminal tensor product.

\subsection{Quasi-Compact Opens Boundary T2}

The following tempting construction is not adopted here:
\[
  \varinjlim_{U\in\mathsf K(X)}\mathcal A(U),
  \qquad
  \mathsf K(X)=\{\text{quasi-compact Zariski opens of }X\}.
\]
For an affine scheme \(X=\operatorname{Spec}R\), the Stacks source cited above
records that \(X\) is quasi-compact.  Hence \(X\in\mathsf K(X)\), and it is a
terminal object under inclusion.  Whenever all the algebras \(\mathcal A(U)\)
are already defined, this colimit is just the global algebra \(\mathcal A(X)\).
That terminal-object fact is harmless in the finite-spectrum examples below,
but it does not model finite-region quasi-local behavior on an infinite affine
space.

Therefore quasi-compact opens may be used here only as topological control data
or in finite/global examples.  They are not the general affine region poset for
the programme until an additional nonterminal region convention, infinite
tensor-product convention, or \(C^*\)-completion convention is fixed.

\subsection{Closed Supports and AQM-28 T3}

The closed-complement assignment from AQM-24 still sends an open \(U\) to
\[
  \mathcal A(X\setminus U).
\]
With opens ordered by inclusion, this assignment is contravariant.  Therefore
it is not the primary open-local system if ``local region grows'' is meant to
match \(U\subset V\).

In finite spectra, the same tensor-factor construction becomes covariant if it
is indexed directly by finite supports, and by finite closed supports when the
intended support is a vanishing locus.  If \(Z,Z'\) are closed subsets of the
finite \(X\), then \(Z\cup Z'\) is closed and finite, and
\(Z,Z'\subset Z\cup Z'\).  The assignment
\[
  Z\longmapsto\mathcal A(Z)
\]
with the same tensor-by-identity inclusions is then a directed system.  This is
best read as a closed-support or vanishing-locus quasi-local algebra, not as
the open-local observable algebra.

For infinite affine schemes of finite type over \(\mathbb F_q\), this shard
does not use all points of all quasi-compact opens as Weyl sites.  AQM-28 gives
the current convention: closed finite-residue sites only, finite supports
\(S\Subset U\cap |X|_{\rm cl}\), and
\[
  \mathcal A(U)
  =
  \varinjlim_{S\Subset U\cap |X|_{\rm cl}}\mathcal A(S).
\]
That is the finite closed-support spine for later infinite affine locality.

\subsection{Finite Checks}

For \(X=\operatorname{Spec}\mathbb F_p\), the space has a largest open
\(X\).  Hence the finite-support directed system has a terminal object and
\[
  \mathcal A_{\rm finqloc}(X)=\mathcal A(X)\cong M_p(\mathbb C).
\]

For \(X=\operatorname{Spec}(\mathbb Z/6\mathbb Z)\), all opens are
finite and \(X\) is again terminal.  The finite-spectrum quasi-local algebra is
\[
  \mathcal A_{\rm finqloc}(X)
  =
  M_2(\mathbb C)\otimes M_3(\mathbb C)\cong M_6(\mathbb C).
\]
The proper open-local subalgebras are the two commuting singleton algebras
\[
  M_2(\mathbb C)\otimes\mathbf 1_3,\qquad
  \mathbf 1_2\otimes M_3(\mathbb C),
\]
and their common upper bound is the full open \(X\).

\subsection{Conclusion}

For AQM-24 finite spectra, the algebraic finite-region convention is
\[
  \boxed{\mathcal A_{\rm finqloc}(X)
  =\varinjlim_{S\in\mathsf F(X)}\mathcal A(S)=\mathcal A(X)}
\]
with \(S\) ranging over finite supports and with inclusions given by tensoring
identities.  The quasi-compact-open colimit is not the general affine
finite-region construction, because affine \(X\) itself is quasi-compact and
terminal.  The closed-complement possibility is retained only as a separate
closed-support construction, useful for vanishing loci and defects.  Later
infinite affine locality should cite the AQM-28 finite closed-support system,
not a quasi-compact-open global colimit from this shard.
