% SHARD-ID: AQM-57-DERHAM-CSS-F2-ENTANGLEMENT-CODES
% SHARD-TITLE: Entanglement Inside the \(\mathbb F_2\) de Rham CSS Prototype
% SHARD-SUMMARY: Derives Bell-pair sectors and code parameters for three prototype \(\mathbb F_2\) de Rham CSS examples.
% SHARD-SUMMARY: Shows explicitly that the prototype examples are entangled stabilizer codes but still have distance-one logical operators.
% SHARD-SUMMARY: Includes an 18 auxiliary-qubit fat-rectangle example with two forced Bell pairs and five logical qubits.
% SHARD-KEYWORDS: F2, qubit, de Rham, CSS, stabilizer code, entanglement, Bell pair, distance
\section{Entanglement Inside the \texorpdfstring{\(\mathbb F_2\)}{F2} de Rham CSS Prototype}
\label{sec:derham-css-f2-entanglement-codes}
\subsection{Status and Local Test for Entanglement}
This shard continues the finite \(C^1\)-basis CSS prototype of AQM-54--AQM-56.
Its qubits and Bell sectors are conditional on the displayed prototype
stabilizers; they are not canonical tangent-Weyl consequences of
\(\Spec A\).  It uses the same source anchors:
\path{references/algebraic_geometry/stacks_project_algebra.tex}, lines
33793--33872, 34103--34120, 34310--34340, and 34388--34524 for K\"ahler
differentials and the de Rham complex, AQM-08 for the CSS/stabilizer
dictionary, and \texttt{CONVENTIONS.md} (ao) for the de Rham CSS test.  The new
claims in this shard are direct solutions of the displayed prototype
stabilizer equations.  For a displayed \(C^1\) basis label \(u\), the notation
\(X_u\) and \(Z_u\) means the corresponding support Pauli, and all listed
stabilizer equations below use the added all-\(+1\) phase convention.  The
de Rham calculation itself fixes only the support spaces \(R_X\) and \(R_Z\).
The local entanglement test used below is the following.  If two qubits \(p,q\)
are constrained by
\[
  X_pX_q=+1,\qquad Z_pZ_q=+1,
\]
and no other displayed generator touches them, then their joint state is
\[
  |\Phi^+\rangle_{pq}=\frac{|00\rangle+|11\rangle}{\sqrt2}.
\]
Indeed, \(Z_pZ_q=+1\) removes the \(|01\rangle\) and \(|10\rangle\)
amplitudes, while \(X_pX_q=+1\) makes the \(|00\rangle\) and \(|11\rangle\)
amplitudes equal.  The reduced density matrix on either qubit is
\[
  \operatorname{diag}(1/2,1/2),
\]
so the state is not a product state.  This is the only entanglement criterion
used in the examples.
\subsection{Example C0: The Eight-Qubit Square Fat Point}
Let
\[
  A_0=\mathbb F_2[x,y]/(x^2,y^2).
\]
Use the \(C^1=\Omega^1_{A_0/\mathbb F_2}\) basis
\[
\begin{array}{llll}
e_1=dx,&e_2=x\,dx,&e_3=y\,dx,&e_4=xy\,dx,\\
e_5=dy,&e_6=x\,dy,&e_7=y\,dy,&e_8=xy\,dy.
\end{array}
\]
This is the eight auxiliary-qubit example from AQM-56.  The de Rham CSS images
are
\[
  R_X=\langle e_1,\ e_5,\ e_3+e_6\rangle,\qquad
  R_Z=\langle e_3+e_6,\ e_4,\ e_8\rangle.
\]
Thus the selected stabilizer generators are
\[
  X_1,\qquad X_5,\qquad X_3X_6,\qquad
  Z_3Z_6,\qquad Z_4,\qquad Z_8.
\]
The six generators are independent and commute; independence is visible from
the private supports \(1,5,4,8\) plus the common two-qubit support \(3,6\).
Therefore the code parameters are
\[
  [[8,\ 8-6,\ d]]=[[8,2,d]].
\]
The pair \(3,6\) is forced to be
\[
  |\Phi^+\rangle_{3,6}.
\]
The remaining pinned qubits are
\[
  |+\rangle_1,\quad |+\rangle_5,\quad |0\rangle_4,\quad |0\rangle_8.
\]
Qubits \(2\) and \(7\) are untouched by every generator, so a code basis is
\[
  |+\rangle_1\,|\alpha\rangle_2\,|\Phi^+\rangle_{3,6}\,
  |0\rangle_4\,|+\rangle_5\,|\beta\rangle_7\,|0\rangle_8,
\]
where \(\alpha,\beta\in\{0,1\}\).  The logical operators may be chosen as
\[
  \bar X_1=X_2,\quad \bar Z_1=Z_2,\qquad
  \bar X_2=X_7,\quad \bar Z_2=Z_7.
\]
Hence \(d=1\).  Inside the prototype this is an entangled stabilizer code, but
not an error-correcting code.
\subsection{Example C1: A Ten-Qubit Fat Rectangle}
Let
\[
  A_1=\mathbb F_2[x,y]/(x^3,y^2).
\]
Since \(d(x^3)=x^2dx\) and \(d(y^2)=0\), the \(dx\)-part has coefficients
\(1,x,y,xy\), while the \(dy\)-part has coefficients
\(1,x,x^2,y,xy,x^2y\).  Use
\[
\begin{array}{llll}
a_0=dx,&a_1=x\,dx,&b_0=y\,dx,&b_1=xy\,dx,\\
c_0=dy,&c_1=x\,dy,&c_2=x^2dy,&d_0=y\,dy,\\
d_1=xy\,dy,&d_2=x^2y\,dy.&
\end{array}
\]
There are ten auxiliary prototype qubits.  The \(d_0\)-image is
\[
  d(x)=a_0,\quad d(y)=c_0,\quad
  d(xy)=b_0+c_1,\quad d(x^2y)=c_2,
\]
so
\[
  R_X=\langle a_0,\ c_0,\ b_0+c_1,\ c_2\rangle.
\]
For \(C^2\), write \(\omega=dx\wedge dy\).  The relation \(x^2dx=0\)
implies \(x^2\omega=0\), so the target basis used for the transpose is
\[
  \omega,\quad x\omega,\quad y\omega,\quad xy\omega.
\]
Against this basis, the nonzero \(d_1\)-values are
\[
  d_1(b_0)=\omega,\quad d_1(c_1)=\omega,\quad
  d_1(b_1)=x\omega,\quad d_1(d_1)=y\omega.
\]
The displayed rows have image rank \(3\); the \(xy\omega\) row is zero.
Thus
\[
  R_Z=\langle b_0+c_1,\ b_1,\ d_1\rangle.
\]
The selected stabilizers are
\[
  X_{a_0},\ X_{c_0},\ X_{b_0}X_{c_1},\ X_{c_2},\qquad
  Z_{b_0}Z_{c_1},\ Z_{b_1},\ Z_{d_1}.
\]
Their rank is \(4+3=7\), so the code is
\[
  [[10,\ 10-7,\ d]]=[[10,3,d]].
\]
The pair \(b_0,c_1\) is forced to be a Bell pair.  The pinned single-qubit
states are
\[
  |+\rangle_{a_0},\quad |+\rangle_{c_0},\quad |+\rangle_{c_2},\quad
  |0\rangle_{b_1},\quad |0\rangle_{d_1}.
\]
The untouched qubits \(a_1,d_0,d_2\) give logical operators
\[
  X_{a_1},Z_{a_1},\qquad X_{d_0},Z_{d_0},\qquad X_{d_2},Z_{d_2}.
\]
Therefore \(d=1\).  The prototype code has a real Bell sector but no
single-qubit error protection.
\subsection{Example C2: An Eighteen-Qubit Fat Rectangle}
Let
\[
  A_2=\mathbb F_2[x,y]/(x^5,y^2).
\]
The relation \(d(x^5)=x^4dx\) removes \(x^4dx\) and \(x^4y\,dx\), while
\(d(y^2)=0\).  Use the \(18\) basis one-forms
\[
\begin{array}{llll}
a_i=x^i dx\ (0\le i\le3),&
b_i=x^iy\,dx\ (0\le i\le3),\\
c_i=x^i dy\ (0\le i\le4),&
d_i=x^iy\,dy\ (0\le i\le4).
\end{array}
\]
The nonzero \(d_0\)-images needed for \(R_X\) are
\[
\begin{array}{rclcrcl}
d(x)&=&a_0,&&d(x^3)&=&a_2,\\
d(y)&=&c_0,&&d(xy)&=&b_0+c_1,\\
d(x^2y)&=&c_2,&&d(x^3y)&=&b_2+c_3,\\
d(x^4y)&=&c_4.&&&
\end{array}
\]
Hence
\[
  R_X=\langle a_0,a_2,c_0,c_2,c_4,b_0+c_1,b_2+c_3\rangle.
\]
Again let \(\omega=dx\wedge dy\), with \(x^4\omega=0\).  The target basis used
for the transpose is
\[
  \omega,x\omega,x^2\omega,x^3\omega,
  y\omega,xy\omega,x^2y\omega,x^3y\omega.
\]
Against this basis, the nonzero \(d_1\)-rows are
\[
\begin{array}{rclcrcl}
d_1(b_0)&=&\omega,&&d_1(c_1)&=&\omega,\\
d_1(b_1)&=&x\omega,&&d_1(b_2)&=&x^2\omega,\\
d_1(c_3)&=&x^2\omega,&&d_1(b_3)&=&x^3\omega,\\
d_1(d_1)&=&y\omega,&&d_1(d_3)&=&x^2y\omega.
\end{array}
\]
The displayed row image has rank \(6\); the \(xy\omega\) and \(x^3y\omega\)
rows are zero.
Thus
\[
  R_Z=\langle b_0+c_1,\ b_1,\ b_2+c_3,\ b_3,\ d_1,\ d_3\rangle.
\]
The selected stabilizers are therefore
\[
\begin{array}{l}
X_{a_0},X_{a_2},X_{c_0},X_{c_2},X_{c_4},
X_{b_0}X_{c_1},X_{b_2}X_{c_3},\\
Z_{b_0}Z_{c_1},Z_{b_1},Z_{b_2}Z_{c_3},Z_{b_3},Z_{d_1},Z_{d_3}.
\end{array}
\]
The rank is \(7+6=13\), so this is an
\[
  [[18,\ 18-13,\ d]]=[[18,5,d]]
\]
CSS code.  The two disjoint pairs \(b_0,c_1\) and \(b_2,c_3\) are forced to be
\[
  |\Phi^+\rangle_{b_0,c_1}\otimes|\Phi^+\rangle_{b_2,c_3}.
\]
The untouched qubits
\[
  a_1,\quad a_3,\quad d_0,\quad d_2,\quad d_4
\]
give five weight-one logical \(X/Z\) pairs, so \(d=1\).  This example reaches
the requested \(15+\)-qubit scale and shows the same diagnosis: the de Rham
CSS test naturally creates entangled stabilizer sectors, but these fat
rectangles still carry unprotected differential logicals.
\subsection{Summary}
The three prototype examples have parameters
\[
\begin{array}{c|c|c|c}
\text{ring} & \text{qubits} & \text{Bell sectors} & [[n,k,d]]\\ \hline
\mathbb F_2[x,y]/(x^2,y^2) & 8 & (y\,dx,\ x\,dy) & [[8,2,1]]\\
\mathbb F_2[x,y]/(x^3,y^2) & 10 & (y\,dx,\ x\,dy) & [[10,3,1]]\\
\mathbb F_2[x,y]/(x^5,y^2) & 18 &
  (y\,dx,\ x\,dy),\ (x^2y\,dx,\ x^3dy) & [[18,5,1]]
\end{array}
\]
So the first prototype qubit-code conclusion is mixed.  Entanglement is exactly
derivable from the displayed \(C^1\)-CSS stabilizers.  However, the same
characteristic-two pathology that creates many global one-form qubits also
leaves weight-one logical operators.  These are prototype entangled stabilizer
codes, not canonical tangent-Weyl dynamics and not useful distance codes.
